---
title: "Global Suicide Statistics"
subtitle: "An Analysis of Critical Variables"
author: "Team BSJ - Shannon Houser, Jack McNeilly, Brian Linder"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA, 
                      message = FALSE, warning = FALSE,
                      fig.height=6, fig.width = 9, fig.align = "center")
```

```{r data}
library(tidyverse)
suicide <- read_csv("SuicideData.csv")
```

## Introduction 

As the Duke community grieves the loss of two of our classmates to suicide 
in the last 2 weeks, our team plans to analyze general global suicide 
rates from 1985-2016 in order to see if there are any prevalant factors 
that might contribute to people taking their own lives. The dataset we
have selected compiles data from four distinct datasets that includes 
information on suicides from over 100 different countries throughout the
world.The data compares socio-economic info with suicide rates by country 
and year. The data is sourced from the World Health Organization, 
the World Bank and, the United Nations Development Program.  
Our goal is to examine these different socio-economic, location,
and gender factors to gain insight regarding how the variables of the 
dataset impact increased suicide rates. Each observation corresponds to the 
number of suicides that occurred in a certain country and within a certain 
age and gender group. The variables include country, year, sex, age group, 
count of suicides, population, suicide rate, country-year composite key,
HDI for year, gdp for year, gdp per capita, and generation. 


## Data Analysis Plan

# Summary Statistics


```{r}
top_ten <- suicide %>% 
  group_by(country) %>% 
  summarise(total_suicides = sum(suicides_no)) %>% 
  arrange(desc(total_suicides)) %>%
  slice(1:10) %>% 
  select(country) %>% 
  pull()

suicide %>% 
  group_by(country) %>% 
  summarise(total_suicides = sum(suicides_no)) %>% 
  arrange(desc(total_suicides)) %>% 
  slice(1:10)
```

```{r}
top_ten_rate <- suicide %>% 
  group_by(country) %>% 
  summarise(rate_suicide = mean(`suicides/100k pop`))%>% 
  arrange(desc(rate_suicide)) %>%
  slice(1:10) %>% 
  select(country) %>% 
  pull()

suicide %>% 
  group_by(country) %>% 
  summarise(rate_suicide = mean(`suicides/100k pop`)) %>% 
  arrange(desc(rate_suicide)) %>% 
  slice(1:10)
```



```{r}
total_suicides_sex <- suicide %>% 
  group_by(sex) %>% 
  summarise(suicides_sex = sum(suicides_no))

ggplot(data = total_suicides_sex, mapping = aes(x = sex, y = suicides_sex))+
  geom_col()+
  labs(title = "Males are 3.3 times more likely than females to commit suicide",
       x = "Sex", y = "Count")
```

```{r}
top_ten_over_time <- suicide %>% 
  group_by(country, year) %>% 
  summarise(per_year = sum(suicides_no)) %>% 
  filter(country %in% c(top_ten))

ggplot(data = top_ten_over_time, aes(x = year, y = per_year, color = country))+
  geom_line()+
  labs(title = "Total Suicides in top 10 countries", x = "Year", y="Total Suicides per Year")
```

The response variables we will test are the numbers of suicides and the
number of suicides per 100k people. The explanatory variables we will 
examine are different age groups, sex, years, countries, and the 
socioeconomic status of each country, including their Human Development
Index (HDI),growth domestic product (GDP), and GDP per capita.

In addition to observing each individual explanatory variable's impact on the response variables, we will examine how the following combination of explanatory variables and the corresponding result on the response variables: 

-Age groups faceted by sex 
-HDI with age 
-HDI with sex 
-HDI with sex and age
-GDP with age
-GDP with sex
-GDP with sex and age
-the above combinations faceted by time period (years) 
-the above combinations for Each individual country
-the above combinations for World regions including continents and
sub-regions of each continent


In our analysis of the dataset we plan to use statistical methods and tools 
in R including, linear modeling, regression modeling, a combination of
visualization techniques, and null hypothesis testing. We will use packages including 

