---
title: "Global Suicide Statistics"
subtitle: "An Analysis of Critical Variables"
author: "Team BSJ - Shannon Houser, Jack McNeilly, Brian Linder"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA, 
                      message = FALSE, warning = FALSE,
                      fig.height=6, fig.width = 9, fig.align = "center")
```

```{r data}
library(tidyverse)
suicide <- read_csv("SuicideData.csv")
```

## Introduction 

As the Duke community grieves the loss of two of our classmates to suicide 
in the last 2 weeks, our team plans to analyze general global suicide 
rates from 1985-2016 in order to see if there are any prevalent factors 
that might contribute to people taking their own lives. The dataset we
have selected compiles data from four distinct datasets that includes 
information on suicides from over 100 different countries throughout the
world.The data compares socio-economic info with suicide rates by country 
and year. The data is sourced from the World Health Organization, 
the World Bank and, the United Nations Development Program.  
Our goal is to examine these different socio-economic, location,
and gender factors to gain insight regarding how the variables of the 
dataset impact increased suicide rates. Each observation corresponds to the 
number of suicides that occurred in a certain country and within a certain 
age and gender group. The variables include country, year, sex, age group, 
count of suicides, population, suicide rate, country-year composite key,
HDI for year, gdp for year, gdp per capita, and generation. 


# Data Analysis Plan

## Summary Statistics and Visualizations


```{r top_ten}
top_ten <- suicide %>% 
  group_by(country) %>% 
  summarise(total_suicides = sum(suicides_no)) %>% 
  arrange(desc(total_suicides)) %>%
  slice(1:10) %>% 
  select(country) %>% 
  pull()

suicide %>% 
  group_by(country) %>% 
  summarise(total_suicides = sum(suicides_no)) %>% 
  arrange(desc(total_suicides)) %>% 
  slice(1:10)
```
```{r top_ten_overtime}
top_ten_over_time <- suicide %>% 
  group_by(country, year) %>% 
  summarise(per_year = sum(suicides_no)) %>% 
  filter(country %in% c(top_ten))

ggplot(data = top_ten_over_time, aes(x = year, y = per_year, 
                                     color = country))+
  geom_line()+
  labs(title = "Total Suicides in top 10 countries", 
       x = "Year", y="Total Suicides per Year", color = "Country")
```
The top ten countries in terms of total suicides seem to be those 
that are most populated. We then decided to graph how the total 
number of suicides in these countries has changed over the years. 
We decided that this is not really helpful information and went on to explore further.

```{r top_ten_rate}
top_ten_rate <- suicide %>% 
  group_by(country) %>% 
  summarise(rate_suicide = mean(`suicides/100k pop`))%>% 
  arrange(desc(rate_suicide)) %>%
  slice(1:10) %>% 
  select(country) %>% 
  pull()

suicide %>% 
  group_by(country) %>% 
  summarise(rate_suicide = mean(`suicides/100k pop`)) %>% 
  arrange(desc(rate_suicide)) %>% 
  slice(1:10)
```

When we examined the suicide rate (per 100k people), we found a strong
correlation between geopolitical circumstances as 9/10 of the top 10 countries
for suicide rate were part of the ex- Soviet Union. They are all Eastern 
European countries that may share history, religions, wars, etc. that we are
unable to currently predict. 

```{r top_ten_rate_over_time}
top_ten_rate_over_time <- suicide %>% 
  group_by(country, year) %>% 
  summarise(rate_suicide = mean(`suicides/100k pop`)) %>% 
  filter(country %in% c(top_ten_rate))

ggplot(data = top_ten_rate_over_time, aes(x = year, rate_suicide, 
                                     color = country))+
  geom_line()+
  labs(title = 
         "Suicide rates spike in the late 90's and continue to decrease",
       subtitle = 
         "Showing 10 countries with highest suicide rates from 1985-2016",
       x = "Year", y="Suicide Rate", color = "Country")
```

Graphing the average suicide rates over time in the ten countries with the
highest suicide rates, we can see that the suicide rates peaked in the late
1990's and have continued to decrease since. 


```{r total_sex}
total_suicides_sex <- suicide %>% 
  group_by(sex) %>% 
  summarise(suicides_sex = sum(suicides_no))

ggplot(data = total_suicides_sex, mapping = aes(x = sex, y = suicides_sex, 
                                                fill = sex))+
  geom_col()+
  scale_fill_manual(values = c("grey 50", "red"))+
  labs(title = 
         "Males are 3.3 times more likely than females to commit suicide",
       x = "Sex", y = "Count")
    
```

From this data visualization, it is obvious that sex is probably a very
important variable when predicting suicide. Men are more than 3.3 times more
likely than women to commit suicide. 

```{r total_age}
total_suicides_age <- suicide %>% 
  group_by(age) %>% 
  summarise(suicides_age = sum(suicides_no))

ggplot(data = total_suicides_age, mapping = aes(x = age, y = suicides_age, 
                                                fill = age))+
  geom_col()+
  scale_x_discrete(limits=c("5-14 years","15-24 years","25-34 years",
                            "35-54 years","55-74 years","75+ years"))+
  scale_fill_manual(values = c("grey 50","grey 50","red", "grey 50",
                               "grey 50","grey 50"))+
  labs(title = 
         "Those Age 35-54 are most at risk for suicide",
       x = "Age Group", y = "Count")
```

This visualization shows us that age may also have a strong influence on the
likeliness of someone to commit suicide. From the last two visualizations, we
may suggest the possibility that middle-aged men are most at risk for suicide.
We may also want to examine this relationship more and see brainstorm what life
factors make middle-aged men more likey than any other group to commit suicide. 

```{r total_gen}
suicide %>% 
  group_by(generation) %>% 
  summarise(tot_gen = sum(suicides_no)) %>% 
  arrange(desc(tot_gen))
```

This shows the total number of suicides per generation. This has a lot to do
with age group and thus is redundant; however, it may help us to better
understand what kinds of life circumstances outside of the data these people 
may have faced to lead them to commit suicide. 

```{r age_over_time}
us_age_time <- suicide %>% 
  group_by(country, age, year) %>% 
  filter(country == "United States") %>% 
  summarise(per_age = sum(suicides_no))

ggplot(data = us_age_time, aes(x = year, y = per_age, color = age))+
  geom_line()+
  labs(title = "Increasing number of Suicides in 35-74 year olds over time", 
       x = "Year", y="Total Suicides per Year", color = "Age")
```

From this visualization we can see that over time, the number of global 
suicides for those between the ages of 35- 74 have increased the most
drastically. The other age groups seem to be roughly stable; however, it does
appear that all other age groups are increasing at the very end of the graph.
It would be interesting to see if this sad trend continued past the last 
year of this study's data collection, 2016. 

```{r gdp_top_ten}
suicide %>%
  group_by(country) %>% 
  summarise(avg_gdp = mean(`gdp_for_year ($)`)) %>% 
  arrange(desc(avg_gdp)) 

top_10_gdp <- suicide %>%
  group_by(country) %>% 
  summarise(avg_gdp = mean(`gdp_for_year ($)`),
            tot_suicides = sum(suicides_no)) %>% 
  arrange(desc(avg_gdp)) 

suicide %>%
  group_by(country) %>% 
  summarise(avg_gdp_capita = mean(`gdp_per_capita ($)`)) %>% 
  arrange(desc(avg_gdp_capita)) 

top_10_gdp_capita <- suicide %>%
  group_by(country) %>% 
  summarise(avg_gdp_capita = mean(`gdp_per_capita ($)`),
            tot_suicides = sum(suicides_no)) %>% 
  arrange(desc(avg_gdp_capita))
```


```{r gdp_vs_total}
ggplot(data = top_10_gdp_capita, mapping = aes(x = avg_gdp_capita,
                                     y = tot_suicides)) +
  geom_point()+
  labs(title = "Minimal relationship between gdp per capita and suicides", 
       x = "GDP per capita", y = "Number of Suicides")
```


## Planning

The response variables we will test are the numbers of suicides and the
number of suicides per 100k people. The explanatory variables we will 
examine are different age groups, sex, years, countries, and the 
socioeconomic status of each country, including their Human Development
Index (HDI),growth domestic product (GDP), and GDP per capita.

In addition to observing each individual explanatory variable's 
impact on the response variables, we will examine how the following 
combination of explanatory variables and the corresponding result 
on the response variables: 

-Age groups faceted by sex 
-HDI with age 
-HDI with sex 
-HDI with sex and age
-GDP with age
-GDP with sex
-GDP with sex and age
-the above combinations faceted by time period (years) 
-the above combinations for Each individual country
-the above combinations for World regions including continents and
sub-regions of each continent


In our analysis of the dataset we plan to use statistical methods and tools 
in R including, linear modeling, regression modeling, a combination of
visualization techniques, and null hypothesis testing. 

In our analysis of the dataset we plan to use statistical methods and tools 
in R including, linear modeling, regression modeling, a combination of
visualization techniques, and null hypothesis testing.

From our preliminary analysis, we believe that trying to find predictors of
suicide using modeling techniques would be a good place to start. We believe 
that such variables as sex and age may have large impacts on the response
variables of total suicide numbers and mean suicide rates. We also believe
that gdp and gdp per capita may not play as large of a role as people may think.
Instead, perhaps geopolitical factors that are outside of our datasets scope
play a large part in suicide determinants. 

Furthermore, we plan to explore how these factors have changed over time
and if the changes are statistically significant. For example, we will explore
whether the total number of suicides in the US has significantly changed between
1985 and 2016, and compare these changes with comparable nations within the data.


# Glimpse

```{r glimpse}
glimpse(suicide)
```


