---
title: "Global Suicide Statistics"
subtitle: "An analysis of critical variables"
author: "Team BSJ - Shannon Houser, Jack McNeilly, Brian Linder"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA, 
                      message = FALSE, warning = FALSE,
                      fig.height=6, fig.width = 9, fig.align = "center")
```

```{r load_packages}
library(tidyverse)
library(rvest)
library(magrittr)
library(ggmap)
library(stringr)
library(maps)
```

```{r clean_names}
library(janitor)
suicide <- janitor::clean_names(suicide)
```

```{r data}
suicide <- read_csv("Data/SuicideData.csv")
```

# Introduction 

As the Duke community grieves the loss of two of our classmates to suicide 
in the last 2 weeks, our team plans to analyze general global suicide 
rates from 1985-2016 in order to see if there are any prevalent factors 
that might contribute to people taking their own lives. The dataset we
have selected compiles data from four distinct datasets that includes 
information on suicides from over 100 different countries throughout the
world.The data compares socio-economic info with suicide rates by country 
and year. The data is sourced from the World Health Organization, 
the World Bank and, the United Nations Development Program.  
Our goal is to examine these different socio-economic, location,
and gender factors to gain insight regarding how the variables of the 
dataset impact increased suicide rates. Each observation corresponds to the 
number of suicides that occurred in a certain country and within a certain 
age and gender group. The variables include country, year, sex, age group, 
count of suicides, population, suicide rate, country-year composite key,
HDI for year, gdp for year, gdp per capita, and generation. 


# Data Analysis Plan

# Planning

The response variables we will test are the numbers of suicides and the
number of suicides per 100k people. The explanatory variables we will 
examine are different age groups, sex, years, countries, and the 
socioeconomic status of each country, including their Human Development
Index (HDI),growth domestic product (GDP), and GDP per capita.

In addition to observing each individual explanatory variable's 
impact on the response variables, we will examine how the following 
combination of explanatory variables and the corresponding result 
on the response variables: 

-Age groups faceted by sex 
-HDI with age 
-HDI with sex 
-HDI with sex and age
-GDP with age
-GDP with sex
-GDP with sex and age
-the above combinations faceted by time period (years) 
-the above combinations for Each individual country
-the above combinations for World regions including continents and
sub-regions of each continent


In our analysis of the dataset we plan to use statistical methods and tools 
in R including, linear modeling, regression modeling, a combination of
visualization techniques, and null hypothesis testing. 

In our analysis of the dataset we plan to use statistical methods and tools 
in R including, linear modeling, regression modeling, a combination of
visualization techniques, and null hypothesis testing.

From our preliminary analysis, we believe that trying to find predictors of
suicide using modeling techniques would be a good place to start. We believe 
that such variables as sex and age may have large impacts on the response
variables of total suicide numbers and mean suicide rates. We also believe
that gdp and gdp per capita may not play as large of a role as people may 
think.

Instead, perhaps geopolitical factors that are outside of our datasets scope
play a large part in suicide determinants. 

Furthermore, we plan to explore how these factors have changed over time
and if the changes are statistically significant. For example, we will explore
whether the total number of suicides in the US has significantly changed 
between 1985 and 2016, and compare these changes with comparable nations 
within the data.

# Glimpse of Data

```{r glimpse}
glimpse(suicide)
```

## Summary Statistics and Visualizations

```{r top_ten}
top_ten <- suicide %>% 
  group_by(country) %>% 
  summarise(total_suicides = sum(suicides_no)) %>% 
  arrange(desc(total_suicides)) %>%
  slice(1:10) %>% 
  select(country) %>% 
  pull()

suicide %>% 
  group_by(country) %>% 
  summarise(total_suicides = sum(suicides_no)) %>% 
  arrange(desc(total_suicides)) %>% 
  slice(1:10)
```

```{r top_ten_overtime}
top_ten_over_time <- suicide %>% 
  group_by(country, year) %>% 
  summarise(per_year = sum(suicides_no)) %>% 
  filter(country %in% c(top_ten))

ggplot(data = top_ten_over_time, aes(x = year, y = per_year, 
                                     color = country))+
  geom_line()+
  labs(title = "Total Suicides in top 10 countries", 
       x = "Year", y="Total Suicides per Year", color = "Country")
```
The top ten countries in terms of total suicides seem to be those 
that are most populated. We then decided to graph how the total 
number of suicides in these countries has changed over the years. 
We decided that this is not really helpful information and went on to explore 
further.

```{r top_ten_rate}
top_ten_rate <- suicide %>% 
  group_by(country) %>% 
  summarise(rate_suicide = mean(suicides_100k_pop))%>% 
  arrange(desc(rate_suicide)) %>%
  slice(1:10) %>% 
  select(country) %>% 
  pull()

suicide %>% 
  group_by(country) %>% 
  summarise(rate_suicide = mean(suicides_100k_pop)) %>% 
  arrange(desc(rate_suicide)) %>% 
  slice(1:10)
```

When we examined the suicide rate (per 100k people), we found a strong
correlation between geopolitical circumstances as 9/10 of the top 10 countries
for suicide rate were part of the ex- Soviet Union. They are all Eastern 
European countries that may share history, religions, wars, etc. that we are
unable to currently predict. We also wanted to include a graph of this data to
focus on the shared geographical component of these results. 




```{r top_ten_rate_over_time}
top_ten_rate_over_time <- suicide %>% 
  group_by(country, year) %>% 
  summarise(rate_suicide = mean(suicides_100k_pop)) %>% 
  filter(country %in% c(top_ten_rate))

ggplot(data = top_ten_rate_over_time, aes(x = year, rate_suicide, 
                                     color = country))+
  geom_line()+
  labs(title = 
         "Suicide rates spike in the late 90's and continue to decrease",
       subtitle = 
         "Showing 10 countries with highest suicide rates from 1985-2016",
       x = "Year", y="Suicide Rate", color = "Country")
```

Graphing the average suicide rates over time in the ten countries with the
highest suicide rates, we can see that the suicide rates peaked in the late
1990's and have continued to decrease since. 


```{r total_sex}
total_suicides_sex <- suicide %>% 
  group_by(sex) %>% 
  summarise(suicides_sex = sum(suicides_no))

ggplot(data = total_suicides_sex, mapping = aes(x = sex, y = suicides_sex, 
                                                fill = sex))+
  geom_col()+
  scale_fill_manual(values = c("grey 50", "red"))+
  labs(title = 
         "Males are 3.3 times more likely than females to commit suicide",
       x = "Sex", y = "Count")
    
```

From this data visualization, it is obvious that sex is probably a very
important variable when predicting suicide. Men are more than 3.3 times more
likely than women to commit suicide. 

```{r total_age}
total_suicides_age <- suicide %>% 
  group_by(age) %>% 
  summarise(suicides_age = sum(suicides_no))

ggplot(data = total_suicides_age, mapping = aes(x = age, y = suicides_age, 
                                                fill = age))+
  geom_col()+
  scale_x_discrete(limits=c("5-14 years","15-24 years","25-34 years",
                            "35-54 years","55-74 years","75+ years"))+
  scale_fill_manual(values = c("grey 50","grey 50","red", "grey 50",
                               "grey 50","grey 50"))+
  labs(title = 
         "Those Age 35-54 are most at risk for suicide",
       x = "Age Group", y = "Count")
```

This visualization shows us that age may also have a strong influence on the
likeliness of someone to commit suicide. From the last two visualizations, we
may suggest the possibility that middle-aged men are most at risk for suicide.
We may also want to examine this relationship more and see brainstorm what 
life factors make middle-aged men more likely than any other group to commit 
suicide. 

```{r total_gen}
suicide %>% 
  group_by(generation) %>% 
  summarise(tot_gen = sum(suicides_no)) %>% 
  arrange(desc(tot_gen))
```

This shows the total number of suicides per generation. This has a lot to do
with age group and thus is redundant; however, it may help us to better
understand what kinds of life circumstances outside of the data these people 
may have faced to lead them to commit suicide. 

```{r age_over_time}
us_age_time <- suicide %>% 
  group_by(country, age, year) %>% 
  filter(country == "United States") %>% 
  summarise(per_age = sum(suicides_no))

ggplot(data = us_age_time, aes(x = year, y = per_age, color = age))+
  geom_line()+
  labs(title = "Increasing number of Suicides in 35-74 year olds over time", 
       x = "Year", y="Total Suicides per Year", color = "Age")
```

From this visualization we can see that over time, the number of global 
suicides for those between the ages of 35- 74 have increased the most
drastically. The other age groups seem to be roughly stable; however, it does
appear that all other age groups are increasing at the very end of the graph.
It would be interesting to see if this sad trend continued past the last 
year of this study's data collection, 2016. 

```{r gdp_top_ten}
suicide %>%
  group_by(country) %>% 
  summarise(avg_gdp = mean(gdp_for_year)) %>% 
  arrange(desc(avg_gdp)) 

top_10_gdp <- suicide %>%
  group_by(country) %>% 
  summarise(avg_gdp = mean(gdp_for_year),
            tot_suicides = sum(suicides_no)) %>% 
  arrange(desc(avg_gdp)) 

suicide %>%
  group_by(country) %>% 
  summarise(avg_gdp_capita = mean(gdp_per_capita)) %>% 
  arrange(desc(avg_gdp_capita)) 

top_10_gdp_capita <- suicide %>%
  group_by(country) %>% 
  summarise(avg_gdp_capita = mean(gdp_per_capita),
            tot_suicides = sum(suicides_no), 
            suicide_per_capita = mean(suicides_100k_pop)) %>% 
  arrange(desc(avg_gdp_capita))
```


```{r gdp_vs_total}
ggplot(data = top_10_gdp_capita, mapping = aes(x = avg_gdp_capita,
                                     y = tot_suicides)) +
  geom_point()+
  labs(title = "Minimal relationship between gdp per capita and suicides", 
       x = "GDP per capita", y = "Number of Suicides")
```

```{r hdi_analysis}
suicide_hdi <- suicide %>%
  na.omit(suicide)

suicide_hdi %>% 
  group_by(country) %>% 
  summarise(avg_hdi = mean(hdi_for_year)) %>% 
  arrange(desc(avg_hdi)) 

top_hdi <- suicide_hdi %>%
  group_by(country) %>% 
  summarise(avg_hdi = mean(hdi_for_year),
            suicide_per_capita = mean(suicides_100k_pop)) %>% 
  arrange(desc(avg_hdi)) 
```

```{r hdi_vs_suicide_per_100k}
ggplot(data = top_hdi, mapping = aes(x = avg_hdi,
                                     y = suicide_per_capita)) +
  geom_point()+
  labs(title = 
         "Minimal relationship between human development index and suicides", 
       x = "Human Development Index", y = "Suicides per 100k")
```

```{r gdp_groupings}
 top_10_gdp_capita <- top_10_gdp_capita %>% 
  mutate(gdp_groups = case_when(avg_gdp_capita < 2000 ~ "<2000", 
                                
                              avg_gdp_capita >= 2000 & avg_gdp_capita < 5000 
                                ~ "2,000-5,000", 
                              avg_gdp_capita >= 5000 & avg_gdp_capita < 10000
                                ~ "5,000-10,000", 
                              avg_gdp_capita >= 10000 & avg_gdp_capita < 20000
                                ~ "10,000-20,000",
                              avg_gdp_capita >= 20000 & avg_gdp_capita < 35000
                                ~ "20,000-35,000",
                              avg_gdp_capita >= 35000 & avg_gdp_capita < 50000
                                ~ "35,000-50,000",
                                avg_gdp_capita > 50000 ~ ">50000"))
```


```{r gdp_groups_suicide}

ggplot(data = top_10_gdp_capita, mapping = aes(x = gdp_groups, 
                              y = suicide_per_capita, fill = gdp_groups)) +
  geom_col() +
  scale_x_discrete(limits=c("<2000",
                            "2,000-5,000",
                            "5,000-10,000",
                            "10,000-20,000",
                            "20,000-35,000",
                            "35,000-50,000",
                            ">50000")) +
  scale_fill_manual(values = c("grey 50","grey 50","grey 50",
                               "grey 50", "grey 50", "grey 50", "red" )) +
  labs(title = 
         "GDP per Capita presents little influence on suicides per 100k",
       x = "GDP per Capita", y = "Average Suicide per 100k")
```

The visualization above divides countries by levels commonly used to analyze 
GDP per capita. The countries with GDP per capita 5,000-10,000 together 
experienced the highest number of suicides per 100k of any group. However, the
group with the next highest suicide rate was those countries with GDP per 
capita of 2,000-5,000, followed by those with GDP per capita of 10,000-20,000,
indicating a lack of a clear pattern between GDP per capita and suicides per 
100k. This visualization possibly provides insight showing that countires 
with mid-to-low GDP per capita, averaged together, have the highest suicide 
rate. However, it further indicates that countries grouped by HDI is a more 
consistant and insighful visualization when compared with suicides per 100k. 

## Try log Graphs

```{r gdp_graph_log}
ggplot(data = suicide_hdi, mapping = aes(x = log(gdp_per_capita),
                                     y = log(suicides_100k_pop))) +
  geom_point()+
  geom_smooth()
```


```{r hdi_graph_log}
ggplot(data = suicide_hdi, mapping = aes(x = hdi_for_year,
                                     y = log(suicides_100k_pop))) +
  geom_point()+
  geom_smooth()
```

These graphs also show that there is a slight linear relationship between GDP 
per capita and HDI and log(suicide rate). This will be helpful in our model. 


#### Methods and Results

## Map
```{r world_map}
top_ten_rate <- suicide %>% 
  group_by(country) %>% 
  summarise(rate_suicide = mean(suicides_100k_pop))%>% 
  arrange(desc(rate_suicide)) %>%
  mutate(country = if_else(country == "Russian Federation", "Russia",
                           country)) %>% 
  slice(1:10) 

map_world <- map_data("world")

map_world_joined <- left_join(map_world, top_ten_rate, by = 
                                c('region' = 'country'))

map_world_joined <- map_world_joined %>% 
  mutate(fill_flg = ifelse(is.na(rate_suicide),F,T))

ggplot() +
  geom_polygon(data = map_world_joined, aes(x = long, y = lat, 
                              group = group, fill = fill_flg)) +
  scale_fill_manual(values = c("#CCCCCC","#e60000")) +
  labs(title = 
  "Countries with Highest Average Suicide Rate are Mostly ex-Soviet Union") +
  theme(text = element_text(color = "#FFFFFF")
        ,panel.background = element_rect(fill = "#212121")
        ,panel.border = element_blank()
        ,plot.background = element_rect(fill = "#212121")
        ,panel.grid = element_blank()
        ,plot.title = element_text(size = 19)
        ,plot.subtitle = element_text(size = 10)
        ,axis.text = element_blank()
        ,axis.title = element_blank()
        ,axis.ticks = element_blank()
        ,legend.position = "none"
        )
```
We wanted to examine which factors were most strongly correlated to high 
suicide rates throughout the world. We also hoped to create a strong linear 
model with these factors in order to accurately predict suicide rates around 
the globe. In order to determine what these most strongly correlated factors 
were, we started with exploratory data analysis. Originally, we looked at 
countries that had the top number of total suicides in the world, but found 
that this was strongly correlated to population. We decided to examine suicide
rate instead, as a better marker. We explored the 10 countries with the 
highest suicide rates (accounting for differences in population) and were 
surprised to see that 9 out of 10 of these countries were once part of the 
ex-Soviet Union. We then mapped this data and decided that there may be 
geo-political forces that influence suicide rates globally. 



## Age + Sex Graphs
```{r sex_and_age_1}
suicide %>% 
  mutate(sex = if_else(sex == "male", "Male", "Female")) %>% 
  ggplot(mapping = aes(x = age, y = suicides_no, 
                                                fill = sex)) +
  geom_col(position = "dodge") +
  scale_x_discrete(limits=c("5-14 years","15-24 years","25-34 years",
                            "35-54 years","55-74 years","75+ years")) +
  coord_flip() +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(title = 
  "Males Age 35-54 Commit Suicide the Most and Males are 3.3 times more likely
than Females to Commit Suicide at any Age",
       x = "Age Group", y = "Suicide count", fill = "Sex") 

```

```{r sex_and_age}
suicide %>% 
  mutate(sex = if_else(sex == "male", "Male", "Female")) %>% 
  ggplot(mapping = aes(x = age, y = suicides_100k_pop, 
                                                fill = sex)) +
  geom_col(position = "dodge") +
  scale_x_discrete(limits=c("5-14 years","15-24 years","25-34 years",
                            "35-54 years","55-74 years","75+ years")) +
  coord_flip() +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(title = 
  "Males age 75+ experience the highest rate of suicide",
       x = "Age Group", y = "Suicides per 100k", fill = "Sex") 

```

We also found that the number of suicides differed greatly among the two 
biological sexes, with men committing suicide 3.3 times more than their female
counterparts. We also explored how the number of total suicides  differed 
among different age groups. We found that the highest total suicides were of 
those aged 35-54. We postulated that this was due to the “mid-life crisis” 
that many face as they deal with growing family and occupational 
responsibilities. We then determined that these two categorical variables 
would be very important and key components of our future linear model. We 
wanted to examine the relationship between these variables and suicide rate. 
When we did this, we saw that those older than 75 had the highest suicide 
rates for both genders. This revealed the sad truth that while middle aged 
adults commit the most suicides in number, the elderly population commits the 
highest proportion of suicides. 

## HDI

```{r hdi_groupings}
top_hdi <- top_hdi %>% 
  mutate(hdi_groups = case_when(
                  avg_hdi >= 0.350 & avg_hdi < 0.550 ~ "low(0.350-0.550)", 
                  avg_hdi >= 0.550 & avg_hdi < 0.700 ~ "medium (0.550-0.700)",
                  avg_hdi >= 0.700 & avg_hdi < 0.800 ~ "high (0.700-0.800)", 
                  avg_hdi >= 0.800 & avg_hdi < 1 ~ "very high (0.800-1)"))
```

```{r hdi_groups_suicide}

ggplot(data = top_hdi, mapping = aes(x = hdi_groups, y = suicide_per_capita, 
                                     fill = hdi_groups))+
  geom_col()+
  scale_x_discrete(limits=c("medium (0.550-0.700)",
                            "high (0.700-0.800)",
                            "very high (0.800-1)"))+
  scale_fill_manual(values = c("grey 50","grey 50","red"))+
  labs(title = 
         "Countries with highest HDI experience higher rate of suicide",
       x = "HDI", y = "Average Suicide per 100k")
```

The next thing that we decided to test was the effect of HDI on global suicide
rates. What we found was that the more developed a country was, the higher 
that country’s rate of suicides were. What this means is that more 
economically and socially advanced countries experience higher rates of these 
tragedies, while their less developed counterparts experience lower rates of 
suicides globally. This is fascinating in the study of socio-economic factors 
on suicide ratees, as it shows that a more developed country does not 
alleviate suicides, and conversely helps their proliferation in these 
countries. 

The visualization above examines a relationship between the Human Development 
Index (HDI) and suicides per 100k. HDI examines a combination of facters 
regarding development, which include life expectancy, education level, and 
gross national income.  We grouped countries into four levels according to 
their assigned HDI value. The categories were as follows: 
-low = 0.350 - 0.550
-medium = 0.550-0.700
-high = 0.700-0.800
-very high = 0.800-1
The visualization reveals that, as a group, countries in the "very high"
category experience the highest number of suicides per 100k. This is followed 
by a comparably high number of sucides per 100k in the "high" category and a 
much lower suicide rate for countries in the "medium" category. This is an 
insighful finding allowing us to understand that the more developed nations on
avaerage suffer from higher rates of suicide, even if their life expectancy (a
componet of HDI) is greater. 


## Linear Model

```{r lm}
library(broom)
lm_total <- lm(suicides_100k_pop ~ factor(sex) + factor(age) + hdi_for_year 
               + gdp_per_capita, data = suicide)

lm_step <- step(lm_total, direction = "backward")

tidy(lm_total)

glance(lm_total) %>% 
  pull(adj.r.squared)
```


```{r check_linearity}
augment(lm_total) %>% 
  ggplot(aes(x = .fitted, y = .resid))+
  geom_point()+
  labs(x = "Predicted Suicide Rate per 100k", y = "Residual Value")+
  geom_hline(yintercept = 0, color = "red", lty = 2)

augment(lm_total) %>% 
  ggplot(aes(x = .resid))+
  geom_histogram(bins = 17)+
  labs(x = "Residual Value", y = "Count")

augment(lm_total) %>% 
  ggplot(aes(x = 1:8364, y = .resid))+
  geom_point()+
  labs(x = "Index", y = "Residual Value")
```

We wanted to fit a linear model to decide which factors were the best when 
predicting suicide rates across the world. We created a linear model and used 
the step function to see which were the best variables for predicting suicide 
rates. However, our first attempt at a linear model was unsuccessful. We saw 
that the data was largely skewed to the right and thought it best to do a data
transformation and predict the log of the suicide rate instead. 

## Log Linear Model
```{r filter_for_model}
library(lubridate)
suicide_hdi <- suicide %>% 
  filter(suicides_100k_pop != 0) %>% 
  mutate(year = ymd(year, truncated = 2L))
```


```{r log_model}
lm_log <- lm(log(suicides_100k_pop) ~ factor(sex) + factor(age) + year
           + hdi_for_year + log(gdp_per_capita), data = suicide_hdi)

lm_step <- step(lm_log, direction = "backward")

tidy(lm_log)

glance(lm_log)
```
$$\widehat{log(suicides/100k)} = sex + age group + year + hdi + log(gdp/capita)$$

This model is a much better predictor. The adjusted R^2 value is .6396, which 
means that the model accurately predicts about 64% of the variation in the log
suicide rate globally. Note that this model was created by data that had to 
filter out the 0s in order to create a better log model. In this way, 
countries with suicide rates of 0 in a particular year are not able to be 
accurately predicted using this model. We can interpret all of the 
coefficients in the model. For example, men will have a log suicide rate that 
is 1.17 higher than their female counterparts, on average. The model also 
suggests, that on average the log rate of suicides has decreases slightly over
time because the coefficient of the year is negative. The model as is, is 
based on females age 15-24, however, we can use the coefficients given to 
change the model to predict the suicide rate of indivisuals with very 
different characteristics. 

We also wanted numeric variables in our model. While  did not see a strong 
relationship between GDP per capita and suicide rate, we still decided to 
include it in our inital linear model. GDP and HDI are our source of 
socio-economic datapoints within this study, so we felt it was important to 
consider their conclusion in the discussion of global suicide rates over time.
We did see that countries with the highest GDPs per capita and HDIs also had 
the highest suicide rates. We also included year because it is an important 
explanatory variable in our dataset, which ranges from 1985-2015. We know that
this time series data breaks the rule of independence when creating a linear 
model. We then created a linear model in order to predict suicide rate based 
on these 5 variables. We saw that the residuals were skewed to the right and 
decided to create a new linear model to predict the log of the suicide rate. 
In order to do this, we filtered out portions of our dataset where the 
observed suicide rate was 0. This made our model much more accurate in 
predicted suicide rates over 0; however, it must be stated that our model 
would not be very good at predicting when it came to countries who have 
suicide rates of 0, and using it to do so would be an extrapolation of the 
model. This log model gave us an R^2 of .6396. This means that are model 
accounts for approximately 64% of the variation in the log suicide rates of 
different countries.


## Linearity Test

```{r check_linearity_2}
augment(lm_log) %>% 
  ggplot(aes(x = .fitted, y = .resid)) +
  geom_point() +
  labs(x = "Predicted log Suicide Rate per 100k", y = "Residual Value") +
  geom_hline(yintercept = 0, color = "red", lty = 2)

augment(lm_log) %>% 
  ggplot(aes(x = .resid)) +
  geom_histogram(bins = 17) +
  labs(x = "Residual Value", y = "Count")

augment(lm_log) %>% 
  ggplot(aes(x = 1:7207, y = .resid)) +
  geom_point() +
  labs(x = "Index", y = "Residual Value")
```

When checking linearity with the log model, it is obvious that the residuals 
are not evenly spread out around 0 at every portion of the line. The residuals
do make an approximately normal distribution. The residuals also do not look 
completely independent of one another; however, this was not unexpected as the
data is time series and therefore the suicide rates are not completely 
independent. 

Furthermore, while the model still does not perfectly pass the linearity 
requirements needed to use this model, it looks much better as a log 
transformation than the original. The residuals have a pretty stable variance 
around 0, but are not present around every portion of the line. Also, the 
residuals, under the log transformation, now form an approximately normal 
distribution. Again, the data is not completely independent because of the 
time series component of the data; however it looks pretty random and 
relatively independent when graphed. In this way, the initial linear 
assumptions that we made in order to form this model are not met completely; 
however, they are only minorly broken, and do not completely discredit the 
model. 


#### Discussion

We chose to work on the global suicide rates dataset to help discuss, and 
uncover some of the driving factors behind suicides over time, across the 
world. In light of the tragedies on Duke’s campus over the last few weeks, we 
felt that it was important to analyze a few of the factors that are most 
prevalant in the discussion over suicide rates. Our goal was to examine 
socio-economic, location, and personal factors, including age and gender, 
to help gain insight into which populations are most susceptible to committing 
suicide. What we found in our analysis is that the largest number of total 
suicides in this dataset have been committed by males, and in particular, males
between the ages of 35-54. This was particularly interesting, as we examine the
factors that impact suicide, this data showed that males are 3.3 times more 
likely than females to commit suicide, and the vast majority of these cases 
are committed by middle aged adults. When we examined the same factors of age 
and gender on suicide rate, we saw that the elderly (defined as those over 75)
committed suicide at the highest rates. This reveals much about society and our
beliefs. We often hear about suicides of middle aged men, which we attribute to
the stress of growing family, financial, and occupation responsibilities. We 
hear about these tragedies in the news because they are the most common in 
number; however, discussion surrounding the suicides of the elderly population 
is less prevalent. While the total number of suicides committed by the elderly 
population cannot compare to those of middle aged adults, they are committing 
the highest proportion of suicides. The mental health struggle of the elderly 
population is thus being minimized because of their smaller population size. 
We also found in this dataset that countries with a higher human development 
index (HDI), a representation of economic and societal growth, experience a 
higher rate of suicide versus less developed countries. This reveals that some
of the most advanced and stable countries in the world are more susceptible to 
suicides from their citizenry. We also were able to create a linear model. The
linear model uses sex, age, year, HDI, and GDP per capita to account for 
approximately 64% of the variance in log suicide rates globally. 

While these findings are fascinating to evaluate, it is important to remember 
that this is just a representative sample of suicide data. Much of the 
information gathered in this study is taken over a long period of time, with 
asymmetrical data available from the different countries within this dataset. 
It is difficult to ensure consistency of reporting of suicides over the 36 year 
study in every country, which could potentially lead to misleading or even 
inaccurate information (for some countries) within our study. Furthermore, as
suicide is a stigmatized topic, many countries may be resistent in reporting 
transparent and fully accurate information, which also leads to potenial 
inconistancies within the dataset.

If given the opportunity to begin the project with a fresh perspective, we 
would have tried to add more data from similar studies in order to increase 
the number of observations, and decrease the potential variance we find in our
results. We would have perferred to find more data on some socio-economic 
relationships with suicide, had we had more explanatory variables to explore 
in the dataset with greater consistency in our results and in the observations
we gathered. We would have preferred to add some more political and social 
factors to use as explanatory variables such as democracy scores, freedom 
indices, happiness scales, or mental health markers. Additionally, our linear
model did not satisfy all the linearity assumptions, so moving forward we would
consider conducting several hypothesis tests to determine if correlation exists
between the given explanatory variables and the suicide rate and exmine the 
nature of that correlation. Given the opportunity to continue this study, we 
would focus specifically on the suicide rates of college-aged students, and 
their suicide rates based on factors that influence their lives as students. 
Trying to bring awareness to the topic of suicide was one of our main goals, 
so had we chosen to continue our analysis, we would have found more 
relationships and data that was pertinent to the colloge student population 
in particular. We would have also focused in on the suicide rates in more
recent years and explored which factors were leading the suicide rates of 
college students to change during this time period. 


# References

Coor_flip()
https://ggplot2.tidyverse.org/reference/coord_flip.html

Converting to date 
https://stackoverflow.com/questions/30255833/convert-four-digit-year-values-to-a-date-type

Map
https://www.r-bloggers.com/how-to-make-a-global-map-in-r-step-by-step/


